Embedded Meta-Learning: Toward more flexible deep-learning models

### 2019-08-11 writing/embedded_meta_learning_main.tex
Incorporating response figure, some clarifications

\tikzstyle{block} = [rectangle, draw, thick, align=center, rounded corners]
\tikzstyle{boundingbox} = [very thick, dotted, gray]
\tikzstyle{dashblock} = [rectangle, draw, thick, align=center, dashed]
\tikzstyle{conc} = [ellipse, draw, thick, dashed, align=center]
\tikzstyle{netnode} = [circle, draw, very thick, inner sep=0pt, minimum size=0.5cm]
\tikzstyle{relunode} = [rectangle, draw, very thick, inner sep=0pt, minimum size=0.5cm]
\tikzstyle{line} = [draw, very thick, -latex']

@@

\vspace{-0.75em} % section
\section{Introduction}
\vspace{-0.5em} % section
Humans are able to use and reuse knowledge more flexibly than most deep learning models can \citep[e.g.][]{Lake2016, Marcus2018}. One fundamental reason for this is that humans are aware of what we are trying to compute and why. By contrast, there is a fundamental separation of knowledge within most deep neural networks -- although deep networks represent knowledge about data (in their activations) and knowledge about the structure of tasks (in their parameters), they do not represent any relationships between data and tasks. That is, a neural network's knowledge about what is being computed is only implicitly accessible to those computations. \par
\looseness=-1
There are a number of advantages to representing knowledge about data and tasks together. In particular, it can grant the ability to rapidly adapt behavior to a new task. The problem of rapid learning has been partially addressed by meta-learning systems \citep[e.g.][see also section \ref{sec_discussion}]{Santoro2016, Finn2017a, Finn2018, Stadie2018, Botvinick2019}. However, humans can use our knowledge of a task to flexibly alter our behavior in accordance with a change in task demands or a single instruction. For example, once we learn to play a game, we can immediately switch to playing in order to lose, and can achieve reasonable performance without any retraining (i.e. zero-shot). Deep learning systems at present generally lack this representational flexibility. \par
In this paper, we propose a new class of tasks based on this idea: meta-mappings, i.e. mappings between tasks (see below). This type of transfer is easily accessible to humans \citep{Lake2016}, but is generally inaccessible to most deep-learning models. To address this challenge, we propose a new class of architectures which essentially take a functional perspective on meta-learning, and exploit the idea of homoiconicity. (A homoiconic programming language is one in which programs in the language can be manipulated by programs in the language, just as data can.) By treating both data and task behaviors as functions, we can conceptually think of both data \emph{and} learned task behaviors as transformable. This yields the ability to not only learn to solve new tasks, but to learn how to transform these solutions in response to changing task demands. We demonstrate that our architectures can flexibly remap their behavior to address the meta-mapping challenge. Furthermore, the architecture may also offer other benefits, such as a useful framework for continual learning. We suggest that approaches like EML will be key to building more intelligent and flexible deep learning systems. \par

@@

\section{Embedded Meta-Learning (EML) architecture}
\vspace{-0.5em} % section
To address these challenges, we propose EML architectures, which are composed of two components: 
\vspace{-1em}
\begin{enumerate} \setlength \itemsep{0em}
\item An input/output system: domain specific encoders and decoders (vision, language, etc.) that map into a shared embedding space $Z$.
\item A meta-learning system which embeds tasks into the shared embedding space $Z$, and learns to map task embeddings to behavior.
\end{enumerate}
\vspace{-0.75em}
Why have a completely shared space $Z$ for data and tasks? The primary advantage is that it allows for arbitrary mappings between these entities. In addition to basic tasks, the system could in principle learn to perform meta-mappings to follow instructions or change behavior. \par
Without training on meta-mappings, of course, the system will not be able to execute them well. However, if it is trained on a broad enough set of such mappings, it will be able to generalize to new instances drawn from the same meta-task distribution, as we demonstrate below. For instances that fall outside its data distribution, or for optimal performance, it may require some retraining, however. This reflects the structure of human behavior -- we are able to learn rapidly when new knowledge is relatively consistent with our previous knowledge, but learning an entirely new paradigm (such as calculus for a new student) can be quite slow \citep[cf.][]{Kumaran2016, Botvinick2019}. \par 
\begin{figure}
\centering
\begin{tikzpicture}[auto, scale=0.75]
\node [dashblock] at (0.2, 0) (rep) {\begin{tabular}{c}$\bm{Z}$: \textbf{Shared} \\ \textbf{representation}\end{tabular}};

@@

\node [conc] at (-7, -3.5) (perc) {\textbf{Input}};
%%\node [conc] at (-7.25, -2.5) (lang) {\textbf{Language}};
\node [conc] at (-7.2, 4) (act) {\textbf{Output}};
\node [conc] at (-7.2, 2.8) (tar) {\textbf{Targets}};
\node [block, text width=2cm] at (-4, -2) (IE) {$\bm{\mathcal{I}}$: \textbf{Input encoder}};
%%\node [block, text width=2cm] at (-5, -0.5) (LE) {$\bm{\mathcal{L}}$: \textbf{Lang. Encoder}};
\node [block, text width=2.2cm] at (-4, 3) (OD) {$\bm{\mathcal{O}}$: \textbf{Output decoder}};
\node [block, text width=2cm] at (-4, 1.5) (TE) {$\bm{\mathcal{T}}$: \textbf{Target encoder}};
\path [line] (perc.north) to (IE.west);
%%\path [line] (lang.north) to ([xshift=0.05cm, yshift=0.05cm]LE.south west);
\path [line] (IE.east) to (rep.south west);
%%\path [line] (LE.east) to (rep.west);
\path [line] (rep.north west) to ([xshift=-0.05cm, yshift=0.05cm]OD.south east);
\path [line] (OD.north) to (act.east);
\path [line] (tar.south) to (TE.west);
\path [line] (TE.east) to (rep.north west);

% meta
\draw [boundingbox] (-1.9, 5) rectangle (6, -5); 
\node [text=gray] at (-0.4, 4.7) {\textbf{Meta Learner}};

@@

\node [block] at (4.2, 0) (meta) {\begin{tabular}{c}$\bm{\mathcal{M}}$: \textbf{Meta} \\ \textbf{network}\end{tabular}};
\path [line] (collection.south) to [out=-90, in=-90] (meta.south);
\path [line] (meta.west) to (rep.east);

@@

\node [block] at (4.2, 3) (hyper) {\begin{tabular}{c}$\bm{\mathcal{H}}$: \textbf{Hyper} \\ \textbf{network}\end{tabular}};
\node [block, dash pattern=on 9pt off 2pt] at (0.2, 2) (transform) {\(F: \bm{Z} \rightarrow \bm{Z}\)};

@@

@@

\caption{The EML architecture. Blocks with solid edges denote learnable deep networks, dashed edges represent inputs, outputs, embeddings, etc., and $F$ is a deep network with parameters specified by $\mathcal{H}$. Note that there may be multiple input, output, and target decoders if the system has multiple input modalities, e.g. both language and vision. See appendix \ref{app_model_details} for details.} \label{architecture_fig}

@@

Given this, the task of mapping inputs to outputs can be framed as trying to find a transformation of the representational space that takes the (embedded) inputs from the training set to embeddings that will decode to the target outputs. These transformations are performed by a system with the following components (see Fig. \ref{architecture_fig} for a schematic): $\mathcal{M}: \{(Z, Z), ...\} \rightarrow Z $ -- the meta network, which collapses a dataset of (input embedding, target embedding) pairs to produce a single function embedding. $\mathcal{H}: Z \rightarrow \text{parameters}$ -- the hyper network, which maps a function embedding to parameters. $F: Z \rightarrow Z$ -- the transformation, implemented by a deep network parameterized by $\mathcal{H}$. \par

>>>>

\tikzstyle{block} = [rectangle, draw, thick, align=center, rounded corners]
\tikzstyle{boundingbox} = [very thick, dotted, gray]
\tikzstyle{dashblock} = [rectangle, draw, thick, align=center, dashed]
\tikzstyle{conc} = [ellipse, draw, thick, dashed, align=center]
\tikzstyle{netnode} = [circle, draw, very thick, inner sep=0pt, minimum size=0.5cm]
\tikzstyle{relunode} = [rectangle, draw, very thick, inner sep=0pt, minimum size=0.5cm]
\tikzstyle{line} = [draw, very thick, -latex']
\tikzstyle{arrow} = [draw, ->, very thick]

@@

\vspace{-0.75em} % section
\section{Introduction}
\vspace{-0.5em} % section
Humans are able to use and reuse knowledge more flexibly than most deep learning models can \citep[e.g.][]{Lake2016, Marcus2018}. One fundamental reason for this is that humans are aware of what we are trying to compute and why. By contrast, there is a fundamental separation of knowledge within most deep neural networks -- although deep networks represent knowledge about data (in their activations) and knowledge about the structure of tasks (in their parameters), they do not represent any relationships between data and tasks. That is, a neural network's knowledge about what is being computed is only implicitly accessible to those computations. \par
\looseness=-1
There are a number of advantages to representing knowledge about data and tasks together. In particular, it can grant the ability to rapidly adapt behavior to a new task. The problem of rapid learning has been partially addressed by meta-learning systems \citep[e.g.][see also section \ref{sec_discussion}]{Santoro2016, Finn2017a, Finn2018, Stadie2018, Botvinick2019}. However, humans can use our knowledge of a task to flexibly alter our behavior in accordance with a change in task demands or a single instruction. For example, once we learn to play a game, we can immediately switch to playing in order to lose, and can achieve reasonable performance without any retraining (i.e. zero-shot). Deep learning systems at present generally lack this representational flexibility. \par
In this paper, we propose a new class of tasks based on this idea: meta-mappings, i.e. mappings between tasks (see below). This type of transfer is easily accessible to humans \citep{Lake2016}, but is generally inaccessible to most deep-learning models. To address this challenge, we propose a new class of architectures which essentially take a functional perspective on meta-learning, and exploit the idea of homoiconicity. (A homoiconic programming language is one in which programs in the language can be manipulated by programs in the language, just as data can.) By treating both data and task behaviors as functions, we can conceptually think of both data \emph{and} learned task behaviors as transformable. This yields the ability to not only learn to solve new tasks, but to learn how to transform these solutions in response to changing task demands. We demonstrate that our architectures can flexibly remap their behavior to address the meta-mapping challenge. By allowing the network to recurse on its representations, our approach is able to achieve this flexibility parsimoniously, without adding additional parameters beyond what we use for basic meta-learning. Furthermore, the architecture may also offer other benefits, such as a useful framework for continual learning. We suggest that approaches like EML will be key to building more intelligent and flexible deep learning systems. \par

@@

\section{Embedded Meta-Learning (EML) architecture}
\vspace{-0.5em} % section
To address these challenges, we propose EML architectures, which are composed of two components: 
\vspace{-1em}
\begin{enumerate} \setlength \itemsep{0em}
\item An input/output system: domain specific encoders and decoders (vision, language, etc.) that map into a shared embedding space $Z$.
\item A meta-learning system which embeds tasks into the shared embedding space $Z$, and learns to map task embeddings to behavior.
\end{enumerate}
\vspace{-0.75em}
Why have a completely shared space $Z$ for data and tasks? The primary advantage is that it allows for arbitrary mappings between these entities. In addition to basic tasks, the system could in principle learn to perform meta-mappings to follow instructions or change behavior. \par
Without training on meta-mappings, of course, the system will not be able to execute them well. However, if it is trained on a broad enough set of such mappings, it will be able to generalize to new instances drawn from the same meta-task distribution, as we demonstrate below. For instances that fall outside its data distribution, or for optimal performance, it may require some retraining, however. This reflects the structure of human behavior -- we are able to learn rapidly when new knowledge is relatively consistent with our previous knowledge, but learning an entirely new paradigm (such as calculus for a new student) can be quite slow \citep[cf.][]{Kumaran2016, Botvinick2019}. \par 
\begin{figure}
\centering
\begin{subfigure}[b]{0.72\textwidth}
\begin{tikzpicture}[auto, scale=0.66, every node/.style={scale=0.85}]
\node [dashblock] at (0.2, 0) (rep) {\begin{tabular}{c}$\bm{Z}$: Shared \\ representation\end{tabular}};

@@

\node [conc] at (-7, -3.5) (perc) {Input};
%%\node [conc] at (-7.25, -2.5) (lang) {Language};
\node [conc] at (-7.2, 4) (act) {Output};
\node [conc] at (-7.2, 2.8) (tar) {Targets};
\node [block, text width=2cm] at (-4, -2) (IE) {$\bm{\mathcal{I}}$: Input encoder};
%%\node [block, text width=2cm] at (-5, -0.5) (LE) {$\bm{\mathcal{L}}$: \textbf{Lang. Encoder}};
\node [block, text width=2.2cm] at (-4, 3) (OD) {$\bm{\mathcal{O}}$: Output decoder};
\node [block, text width=2cm] at (-4, 1.5) (TE) {$\bm{\mathcal{T}}$: Target encoder};
\path [line] (perc.north) to (IE.west);
%%\path [line] (lang.north) to ([xshift=0.05cm, yshift=0.05cm]LE.south west);
\path [line] (IE.east) to (rep.south west);
%%\path [line] (LE.east) to (rep.west);
\path [line] (rep.north west) to ([xshift=-0.05cm, yshift=0.05cm]OD.south east);
\path [line] (OD.north) to (act.east);
\path [line] (tar.south) to (TE.west);
\path [line] (TE.east) to (rep.north west);

% meta
\draw [boundingbox] (-1.9, 5) rectangle (5.85, -5); 
\node [text=gray] at (-0.4, 4.7) {\textbf{Meta Learner}};

@@

\node [block] at (4.2, 0) (meta) {\begin{tabular}{c}$\bm{\mathcal{M}}$: Meta \\ network\end{tabular}};
\path [line] (collection.south) to [out=-90, in=-90] (meta.south);
\path [line] (meta.west) to (rep.east);

@@

\node [block] at (4.2, 3) (hyper) {\begin{tabular}{c}$\bm{\mathcal{H}}$: Hyper \\ network\end{tabular}};
\node [block, dash pattern=on 9pt off 2pt] at (0.2, 2) (transform) {\(F: \bm{Z} \rightarrow \bm{Z}\)};

@@

\end{tikzpicture}
\caption{Schematic of the architecture}
\end{subfigure}%
\begin{subfigure}[b]{0.28\textwidth}
\begin{tikzpicture}[auto, scale=0.66, every node/.style={scale=0.85}]
\node at (-1.5, -3.6) (examples) {Examples};
\node at (-1.5, -2.5) (D1) {
\(\left\{
\begin{matrix}
(z^{ex}_{in,0}, z^{ex}_{targ,0})\\
$\vdots$
\end{matrix}\right\}\)};

\node at (1.5, -3.6) (examples) {Probes};
\node at (1.5, -2.5) (D2) {
%\(z^{prb}_{in}\)};
\(\left\{
\begin{matrix}
z^{prb}_{in,0}\\
$\vdots$
\end{matrix}\right\}\)};

\node at (-0.33, -0.66) (zfunc) {\(z^{func}\)};

\path [arrow] (D1) to [in=-145, out=90] ([xshift=3, yshift=3]zfunc.south west) node [xshift=-16, yshift=-5] {\(\mathcal{M}\)};

\node at (1.5, -0.25) (F) {\(F\)};

\path [arrow, bend left] ([xshift=-5, yshift=-5]zfunc.north east) to [in=165, out=25] ([xshift=3]F.west) node [xshift=-13, yshift=5]{\(\mathcal{H}\)};

\path [arrow, dashed] (D2) to (F);

\node at (1.5, 1.75) (outputs) {
%\(z^{pred}_{out}\)};
\(\left\{
\begin{matrix}
z^{pred}_{out,0},\\
$\vdots$
\end{matrix}\right\}\)};
\node at (1.5, 2.9) (predictions) {Predictions};

\path [arrow, dashed] (F) to ([yshift=3]outputs.south);

\node [align=center, text width=1.25 cm] at (-1.5, 1.75) (dispatch) {Loss or eval or ...};

\path [arrow, dashed] (outputs.west) to ([xshift=-3]dispatch.east);

% for spacing

\node at (0, -5.5) {};

@@

\caption{Meta-learner inference}
\caption{The EML architecture and an inference step. (a) A schematic of the architecture. Blocks with solid edges denote learnable deep networks, dashed edges represent inputs, outputs, embeddings, etc., and $F$ is a deep network with parameters specified by $\mathcal{H}$. Note that there may be multiple input, output, and target decoders if the system has multiple input modalities, e.g. both language and vision. See appendix \ref{app_model_details} for details. (b) Depiction of a meta-learner forward pass. This process is identical between basic tasks and meta-mappings, except for where the example and probe embeddings come from, and how the loss or evaluation is computed. For basic inference, the example and probe embeddings come from basic inputs and outputs, and loss is computed in the output space. FOr meta-mapping, the example and probe input and outputs are task embeddings of basic tasks, and loss for training is computed in the embedding space, while evaluation is done by using the predicted embedding to perform the target task.} \label{architecture_inference_fig}

@@

Given this, the task of mapping inputs to outputs can be framed as trying to find a transformation of the representational space that takes the (embedded) inputs from the training set to embeddings that will decode to the target outputs. These transformations are performed by a system with the following components (see Fig. \ref{architecture_inference_fig}): $\mathcal{M}: \{(Z, Z), ...\} \rightarrow Z $ -- the meta network, which collapses a dataset of (input embedding, target embedding) pairs to produce a single function embedding. $\mathcal{H}: Z \rightarrow \text{parameters}$ -- the hyper network, which maps a function embedding to parameters. $F: Z \rightarrow Z$ -- the transformation, implemented by a deep network parameterized by $\mathcal{H}$. \par
